Realistically Interesting Domains for Probabilistic Planners

### 2014-04-07 FinalPaper/paper.tex
Use textsc for glutton and gourmand. Fix errors (et. all) and other typos

In this paper we evaluate the strength of FF-Replan, winner of IPPC-04 and unenrolled winner of IPPC-06, a replanner that uses the classical planner FF on a determinized probabilistic problem against a state of the art probabilistic planner GOURMAND, the successor of the Labelled Real-Time Dynamic Probabilistic planner Glutton, runner up of the IPPC-11. By translating the Navigation and Traffic domain from IPPC-11 to a matching real-world equivalent and evaluating these using the proposed probabilistically interesting baseline test as proposed by Little et. all we evaluate whether FF-Replan is in fact capable enough to solve enough of these so called classical planning problems in disguise.
\end{abstract}

@@

The IPPC-11 introduced Relation Dynamic Influence Diagram Language (RDDL),
which can design domains that are more probabilistically interesting. The top
performers were PROST \cite{keller2012prost} and Glutton
\cite{kolobov2012glutton}.

@@

The main contribution of this paper is an evaluation of FF-Replan versus a
probabilistically state of the art planner GOURMAND in the realistically
augmented domains of IPPC-11. If these domains fail the baseline test proposed
by Little et\@. all, they would translate to classical planning problems in
disguise and could be easily solved by FF-Replan.

@@

\section{Problem formulation}
To evaluate FF-Replan and GOURMAND in a real world domain a finite horizon MDP is used to represent the realistic \emph{natural dynamics} such as external forces or forces of nature. These dynamics introduce large branching factors emphasizing the need for a planner to be scalable. To evaluate the model, the planner should also be sensitive to an expected reward maximization criterion.
Current finite horizon MDP problems as introduced in IPPC-11 are simplified and therefor do not account for the rich set of state-actions as would be available in a real-world planning problem. We will reason about the most interesting domains for our evaluation and adjust them to match their real-world equivalent. Using the baseline test introduced by Little et. all \cite{little2007probvsreplan}, we can evaluate the probabilistically interestingness of these domains and evaluate the performance of FF-Replan vs. a state of the art probabilistic planner: GOURMAND.

@@

\subsubsection{Probabilistic planning}
Probabilistic planners are used to solve probabilistic planning problems and should be able to account for uncertainty as opposed to classical planners. For our evaluation we have chosen to compare FF-Replan to a state-of-the-art probabilistic planner: GOURMAND. GOURMAND is the online successor of the GLUTTON plan, a Labelled Real-Time Dynamic Programming (LRTDP) based planner and the runner up of IPPC-11.

@@

\subsection{FF-replan}
At the first international probabilistic planning competition IPPC-04, most entries were probabilistic planner using Markov Decision Processes. The winner, FF-Replan \cite{FFReplan}, however was based on deterministic planning techniques. FF-Replan is an action selection algorithm for online planning in probabilistic domains which deals with uncertainty by replanning. FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of the planning. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state mapped in an initially empty hash-table. This plan is executed until an unexpected state, not available in the hash-table is encountered.

@@

FF-Replan has some obvious shortcomings.Iit first determinizes the input domain and thus removes all probabilistic information from the problem. The success of the planner is therefore fully reliable on the chose determinization heuristic. As FF-replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why FF-replan performs poorly in domains containing dead-ends. However, by relying on search FF-replan functions with enormous speed and efficiency and caused the International Conference on Automated Planning and Scheduling (ICAPS) to adapt their domains, increasing their \emph{probabilistically interestingness}. Finite horizon MDPs were introduced as an approximation of the infinite optimization problems using a finite horizon. These planning problems accounted for large branching factors as would be caused by \emph{natural dynamics} such as environmental factors and decay. To be successful in these domains planners need to be scalable and sensitive to an expected reward maximization, the evaluation criteria of finite-horizon MDPs.

@@

\todo{Make this list more technical}
\begin{itemize}
	\item Trials start at $s_0$
	\item Choose best action greedily in the current state based on current $V$ value function
	\item Perform Bellman backup on the current state
	\item Sample the outcome as the chosen action
	\item Transition to the new state
	\item Repeat until a goal, dead end or converged state is reached
	\item After trial, special convergence check on all states in the trial is performed to prove the convergence of these values
\end{itemize}

@@

A \emph{Labeled Reverse Real-Time Dynamic Programmer} (LR\textsuperscript{2}TDP) uses \emph{reverse iterative deepening} and LRTDP-FH to incrementally solve a sequence of MDPs $M(1)$, $M(2)$, \ldots , $M(H)$ in order by relying on the fact that a successful policy can be produced by solving an MDP-FH $M(H)$ for a smaller horizon $h < H$. A LR\textsuperscript{2}TDP\@ iterates on the distance from the horizon and allows for reusing precomputed state-action plans from converged states. However when dealing with large branching factors Bellman computations tend to become prohibitively expensive. 
\todo{nog een lijstje schrijven van hoe het werkt en waarom er verbetering nodig is}

@@

\subsection{GLUTTON}

\textsc{GLUTTON} introduces an off-line algorithm based on LR\textsuperscript{2}TDP. The \emph{Max-Reward} heuristic used to evaluate \todo{the what?} relies on knowing the maximum reward $R_{max}$ any action can yield in a state or an upper bound on it. In addition to LR\textsuperscript{2}TDP, it introduces the following mechanisms to optimize for large branching factors:

\todo{rewrite this list}
\myworries{is this really necessary?}
\begin{itemize}
	\item Sub sampling the transition function, is not enough to decrease the amount of computations yet
	\item Separating out Natural Dynamics: to reuse variable samples when generator successors for multiple actions, by separating the actions' effects into natural dynamics (\emph{exogenous}), those due to the action itself (\emph{pure effects}) and those do a mix of both.
	\item Caching the transition function samples: as LR\textsuperscript{2}TDP populates the state-value hash table lazily, Glutton uses  this memory as a cache for samples from the transition function, so it doesn't have to regenerate the action when encountering the same $(s,h)$ pair. When full, it chooses state-action pars for eviction ans replacement randomly.
	\item Default policies: Glutton uses 3 tricks to determine a policy when stopped prematurely. It compares the optimal action from a $h'$ with a simple cyclic policy or a complete random policy and chooses the best of these based on the reward.
\end{itemize}

%---------------------------------------------------------------------
%During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 GLUTTON was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, GOURMAND is an online adaptation of the GLUTTON algorithm to test whether LRTDP performs better if it is employed online.
%
%UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

Because real world domains contain Natural Dynamics that introduce large branching factors, offline planning is often a waste of time and memory space since there is a vast amount of states the planner may never reach in a lifetime. Such problems are far better solved using online planners by finding an optimal policy $\tau$* or an approximation of this polict for the \emph{current} state, executing it and repeating. This is done by choosing a certain lookahead value $L$ and solving the MDP \emph{M\textsubscript{s}(L)} to solve \emph{M\textsubscript{0}(H)} where L $<$ H and then select \emph{$\pi$*\textsubscript{L}(s, L)} using the evaluation criteria. This lookahead value is determined using the per-process time constraint.\todo{elaborate on the per-process time constraint}

\subsection{GOURMAND}
\textsc{GOURMAND}, an online adaptation of the \textsc{GLUTTON} planner answers these remarks by adding a lookahead value $L$. \todo{still finish this}

% always makes an informed choice in any state where it ends up when the time T allocated for solving the MDP runs out because of the online fashion it uses LRTDP. GOURMAND first divides the time among all remaining epochs, determines the largest lookahead L for which it can almost certainly solve any remaining epoch if T is assigned.  It then checks if it can solve even for a larger lookahead L by taking away a small configuration time from future epochs  while still guaranteeing that it can solve for Lookahead L in each of them. This was GOURMAND adaptively picks a good lookahead value at each decision epoch given the per-process constraint T. This can only be achieved by knowing how long it takes for LRTDP to solve a state for certain lookahead values. It uses LR\textsuperscript{2}TDP to collect this data. This version of LRTDP is more efficient because of the early convergences and it allows measuring how long on average the solving of a task takes. GOURMAND uses the same engineering optimizations to to handle evaluate actions, iterating over the successors of s under a by the Bellman Backups. GOURMAND eventually outperforms both GLUTTON and PROST, the winner of the IPPC-11\cite{keller2012prost}.

@@

\section{Domains}
Following the theoretical comparison of probabilistic planning and replanning techniques by Little and Thi\'ebauw, a baseline test for \emph{probabilistically interesting} domains is proposed. Problems that fail this test were deemed to be classical planning problems in disguise, favouring the use of replanners \cite{little2007probvsreplan}. This section describes interesting domains in which either FF-Replan or GOURMAND prevailed during the International Probabilistic Planning Competitions. These domains are evaluated for how well they portray real-world scenarios. The domains we deem most interesting for the purpose of this paper are adjusted to match their real-world scenario. Little et all's baseline test helps us to predict which planner would perform well on these domains. We will then conclude our evaluation by comparing these results to the result of our experiments described in Section \ref{sec:experiment}.

@@

GOURMAND vastly outperformed IPPC-11 UCT-based probabilistic track winner PROST by far on goal-oriented domains Navigation and Crossing Traffic \cite{kolobov2012gourmand}. PROST however mildly outperforms GOURMAND on the Game of Life, Sysadmin and the Traffic domain as UCT does not need to perform expensive Bellman Backups on these domains that include large branching factors.

@@

The navigation domain represents a grid in which a robot must reach a goal. Every cell offers the robot a different chance of disappearing. The robot needs to choose a path which gets it to the goal most reliably within the finite horizon time \cite{sanner2011navigation}. Because GOURMAND uses a dynamic lookahead, it performs quite well in this domain.

@@

Gourmand already works with RDDLSim and MDPs with finite horizon. The original
FF-Replan is a Goal-Oriented MDP planner. For a Finite-Horizon each state at
the horizon is a goal. To compare Gourmand, that works with MDP-FH, with
FF-Replan, FF-Replan needs a slight update. It should not only execute FF for
one goal, but for all goals at the horizon. It could then start executing the
plan with the highest reward. Depending on the determinization method,
\emph{single-outcomes} or \emph{all-outcomes}, not all goals might be
reachable. Of course the performance decreases with an increase of goal states
at the horizon. A hash-table cache, as FF-Replan already uses can still be used
to store plans generated by FF from a state to goal states, however not only
the plan should be stored, but also the reward for this plan should be stored
to pick the plan with the highest expected reward.

Using the framework of Glutton and Gourmand, and the FF heuristic found in mGPT
\cite{bonet2011mgpt}, the updated FF-Replan can be implemented. The first step
of Gourmand is parsing the problem and creating a structure of the problem.
Instead of executing the Gourmand algorithm, this structure can be used to
determinize the problem by setting all probabilities of each action to 1
(\emph{all-outcome}) or all but one to 0 (\emph{single-outcome}). Using the
FF heuristic it can find plans for each goal and calculate the expected reward.
Then it can reuse the Gourmand framework to execute the plan by sending the
action to the RDDLSim server. Like FF-Replan if the server returns with a state
that is not in the plan the algorithm will replan. If a goal is reached it will
output the achieved reward.

@@

\subsubsection{RFF}. Also uses an incremental plan aggregation, but offline similar to GLUTTON.

>>>>

In this paper we evaluate the strength of FF-Replan, winner of IPPC-04 and unenrolled winner of IPPC-06, a replanner that uses the classical planner FF on a determinized probabilistic problem against a state of the art probabilistic planner \textsc{Gourmand}, the successor of the Labelled Real-Time Dynamic Probabilistic planner \textsc{Glutton}, runner up of the IPPC-11. By translating the Navigation and Traffic domain from IPPC-11 to a matching real-world equivalent and evaluating these using the proposed probabilistically interesting baseline test as proposed by Little et all\. we evaluate whether FF-Replan is in fact capable enough to solve enough of these so called classical planning problems in disguise.
\end{abstract}

@@

The IPPC-11 introduced Relation Dynamic Influence Diagram Language (RDDL),
which can design domains that are more probabilistically interesting. The top
performers were PROST \cite{keller2012prost} and \textsc{Glutton}
\cite{kolobov2012glutton}.

@@

The main contribution of this paper is an evaluation of FF-Replan versus a
probabilistically state of the art planner \textsc{Gourmand} in the realistically
augmented domains of IPPC-11. If these domains fail the baseline test proposed
by Little et all\., they would translate to classical planning problems in
disguise and could be easily solved by FF-Replan.

@@

\section{Problem formulation}

To evaluate FF-Replan and \textsc{Gourmand} in a real world domain a finite horizon MDP is used to represent the realistic \emph{natural dynamics} such as external forces or forces of nature. These dynamics introduce large branching factors emphasizing the need for a planner to be scalable. To evaluate the model, the planner should also be sensitive to an expected reward maximization criterion.
Current finite horizon MDP problems as introduced in IPPC-11 are simplified and therefor do not account for the rich set of state-actions as would be available in a real-world planning problem. We will reason about the most interesting domains for our evaluation and adjust them to match their real-world equivalent. Using the baseline test introduced by Little et all.\@ \cite{little2007probvsreplan}, we can evaluate the probabilistically interestingness of these domains and evaluate the performance of FF-Replan vs.\@ a state of the art probabilistic planner: \textsc{Gourmand}.

@@

\subsubsection{Probabilistic planning}
Probabilistic planners are used to solve probabilistic planning problems and should be able to account for uncertainty as opposed to classical planners. For our evaluation we have chosen to compare FF-Replan to a state-of-the-art probabilistic planner: \textsc{Gourmand}. \textsc{Gourmand} is the online successor of the \textsc{Glutton} plan, a Labelled Real-Time Dynamic Programming (LRTDP) based planner and the runner up of IPPC-11.

@@

\subsection{FF-Replan}
At the first international probabilistic planning competition IPPC-04, most entries were probabilistic planner using Markov Decision Processes. The winner, FF-Replan \cite{FFReplan}, however was based on deterministic planning techniques. FF-Replan is an action selection algorithm for online planning in probabilistic domains which deals with uncertainty by replanning. FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of the planning. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state mapped in an initially empty hash-table. This plan is executed until an unexpected state, not available in the hash-table is encountered.

@@

FF-Replan has some obvious shortcomings. It first determinizes the input domain and thus removes all probabilistic information from the problem. The success of the planner is therefore fully reliable on the chose determinization heuristic. As FF-Replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why FF-Replan performs poorly in domains containing dead-ends. However, by relying on search FF-Replan functions with enormous speed and efficiency. This caused the International Conference on Automated Planning and Scheduling (ICAPS) to adapt their domains, increasing their \emph{probabilistically interestingness}. Finite horizon MDPs were introduced as an approximation of the infinite optimization problems using a finite horizon. These planning problems accounted for large branching factors as would be caused by \emph{natural dynamics} such as environmental factors and decay. To be successful in these domains planners need to be scalable and sensitive to an expected reward maximization, the evaluation criteria of finite-horizon MDPs.

@@

\todo{Make this list more technical}

\begin{itemize}
	\item Trials start at $s_0$
	\item Choose best action greedily in the current state based on current $V$ value function
	\item Perform Bellman backup on the current state
	\item Sample the outcome as the chosen action
	\item Transition to the new state
	\item Repeat until a goal, dead end or converged state is reached
	\item After trial, special convergence check on all states in the trial is performed to prove the convergence of these values
\end{itemize}

@@

A \emph{Labeled Reverse Real-Time Dynamic Programmer} (LR$^2$TDP) uses \emph{reverse iterative deepening} and LRTDP-FH to incrementally solve a sequence of MDPs $M(1)$, $M(2)$, \ldots , $M(H)$ in order by relying on the fact that a successful policy can be produced by solving an MDP-FH $M(H)$ for a smaller horizon $h < H$. A LR$^2$TDP\@ iterates on the distance from the horizon and allows for reusing precomputed state-action plans from converged states. However when dealing with large branching factors Bellman computations tend to become prohibitively expensive.

\todo{nog een lijstje schrijven van hoe het werkt en waarom er verbetering nodig is}

@@

\subsection{\textsc{Glutton}}

\textsc{Glutton} introduces an off-line algorithm based on LR$^2$TDP. The \emph{Max-Reward} heuristic used to evaluate \todo{the what?} relies on knowing the maximum reward $R_{max}$ any action can yield in a state or an upper bound on it. In addition to LR$^2$TDP, it introduces the following mechanisms to optimize for large branching factors:

\todo{rewrite this list}

\myworries{is this really necessary?}

\begin{itemize}
	\item Sub sampling the transition function, is not enough to decrease the amount of computations yet
	\item Separating out Natural Dynamics: to reuse variable samples when generator successors for multiple actions, by separating the actions' effects into natural dynamics (\emph{exogenous}), those due to the action itself (\emph{pure effects}) and those do a mix of both.
	\item Caching the transition function samples: as LR$^2$TDP populates the state-value hash table lazily, \textsc{Glutton} uses  this memory as a cache for samples from the transition function, so it doesn't have to regenerate the action when encountering the same $(s,h)$ pair. When full, it chooses state-action pars for eviction ans replacement randomly.
	\item Default policies: \textsc{Glutton} uses 3 tricks to determine a policy when stopped prematurely. It compares the optimal action from a $h'$ with a simple cyclic policy or a complete random policy and chooses the best of these based on the reward.
\end{itemize}

%---------------------------------------------------------------------
%During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 \textsc{Glutton} was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, \textsc{Gourmand} is an online adaptation of the \textsc{Glutton} algorithm to test whether LRTDP performs better if it is employed online.
%
%UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

Because real world domains contain Natural Dynamics that introduce large branching factors, offline planning is often a waste of time and memory space since there is a vast amount of states the planner may never reach in a lifetime. Such problems are far better solved using online planners by finding an optimal policy $\tau^*$ or an approximation of this policy for the \emph{current} state, executing it and repeating. This is done by choosing a certain lookahead value $L$ and solving the MDP $M_s(L)$ to solve $M_0(H)$ where L $<$ H and then select $\pi^*_L(s,L)$ using the evaluation criteria. This lookahead value is determined using the per-process time constraint.

\todo{elaborate on the per-process time constraint}

\subsection{\textsc{Gourmand}}
\textsc{Gourmand}, an online adaptation of the \textsc{Glutton} planner answers these remarks by adding a lookahead value $L$. \todo{still finish this}

% always makes an informed choice in any state where it ends up when the time T allocated for solving the MDP runs out because of the online fashion it uses LRTDP. \textsc{Gourmand} first divides the time among all remaining epochs, determines the largest lookahead L for which it can almost certainly solve any remaining epoch if T is assigned.  It then checks if it can solve even for a larger lookahead L by taking away a small configuration time from future epochs  while still guaranteeing that it can solve for Lookahead L in each of them. This was \textsc{Gourmand} adaptively picks a good lookahead value at each decision epoch given the per-process constraint T. This can only be achieved by knowing how long it takes for LRTDP to solve a state for certain lookahead values. It uses LR\textsuperscript{2}TDP to collect this data. This version of LRTDP is more efficient because of the early convergences and it allows measuring how long on average the solving of a task takes. \textsc{Gourmand} uses the same engineering optimizations to to handle evaluate actions, iterating over the successors of s under a by the Bellman Backups. \textsc{Gourmand} eventually outperforms both \textsc{Glutton} and PROST, the winner of the IPPC-11\cite{keller2012prost}.

@@

\section{Domains}
Following the theoretical comparison of probabilistic planning and replanning techniques by Little and Thi\'ebauw, a baseline test for \emph{probabilistically interesting} domains is proposed. Problems that fail this test were deemed to be classical planning problems in disguise, favouring the use of replanners \cite{little2007probvsreplan}. This section describes interesting domains in which either FF-Replan or \textsc{Gourmand} prevailed during the International Probabilistic Planning Competitions. These domains are evaluated for how well they portray real-world scenarios. The domains we deem most interesting for the purpose of this paper are adjusted to match their real-world scenario. Little et all's baseline test helps us to predict which planner would perform well on these domains. We will then conclude our evaluation by comparing these results to the result of our experiments described in Section \ref{sec:experiment}.

@@

\textsc{Gourmand} vastly outperformed IPPC-11 UCT-based probabilistic track winner PROST by far on goal-oriented domains Navigation and Crossing Traffic \cite{kolobov2012gourmand}. PROST however mildly outperforms \textsc{Gourmand} on the Game of Life, Sysadmin and the Traffic domain as UCT does not need to perform expensive Bellman Backups on these domains that include large branching factors.

@@

The navigation domain represents a grid in which a robot must reach a goal. Every cell offers the robot a different chance of disappearing. The robot needs to choose a path which gets it to the goal most reliably within the finite horizon time \cite{sanner2011navigation}. Because \textsc{Gourmand} uses a dynamic lookahead, it performs quite well in this domain.

@@

\textsc{Gourmand} already works with RDDLSim and MDPs with finite horizon. The original
FF-Replan is a Goal-Oriented MDP planner. For a Finite-Horizon each state at
the horizon is a goal. To compare \textsc{Gourmand}, that works with MDP-FH, with
FF-Replan, FF-Replan needs a slight update. It should not only execute FF for
one goal, but for all goals at the horizon. It could then start executing the
plan with the highest reward. Depending on the determinization method,
\emph{single-outcomes} or \emph{all-outcomes}, not all goals might be
reachable. Of course the performance decreases with an increase of goal states
at the horizon. A hash-table cache, as FF-Replan already uses can still be used
to store plans generated by FF from a state to goal states, however not only
the plan should be stored, but also the reward for this plan should be stored
to pick the plan with the highest expected reward.

Using the framework of \textsc{Glutton} and \textsc{Gourmand}, and the FF heuristic found in mGPT
\cite{bonet2011mgpt}, the updated FF-Replan can be implemented. The first step
of \textsc{Gourmand} is parsing the problem and creating a structure of the problem.
Instead of executing the \textsc{Gourmand} algorithm, this structure can be used to
determinize the problem by setting all probabilities of each action to 1
(\emph{all-outcome}) or all but one to 0 (\emph{single-outcome}). Using the
FF heuristic it can find plans for each goal and calculate the expected reward.
Then it can reuse the \textsc{Gourmand} framework to execute the plan by sending the
action to the RDDLSim server. Like FF-Replan if the server returns with a state
that is not in the plan the algorithm will replan. If a goal is reached it will
output the achieved reward.

@@

\subsubsection{RFF}. Also uses an incremental plan aggregation, but offline similar to \textsc{Glutton}.
